Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-11 prop_letter.tex
done with sec 2 first pass

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantum Measurements}

@@

The above axioms describe a mathematical idealization involving
infinitely precise measurements. In an actual experimental setup with
an ensemble of quantum states that are not exactly identical, and with
small imperfections and inaccuracies in measuring devices, an
experimenter might only be able to determine that the probability of
the event $\proj{0}$ is in the range $[0.49,0.51]$. The spread in this
range depends on the amount of resources (time, energy, money, etc)
that are devoted to the experiment. In the classical setting, this
``fuzziness'' is formalized by moving to \emph{interval-valued
  probability measures} (IVPMs).

@@

\newpage

Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

In conventional quantum theory (CQT), when we measure a property
$\mathbf{O}$ of a physical system whose state is represented by a
density matrix~$\rho$, the probability to obtain the value $\lambda$
of that property is determined by the Born rule
$\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)=\Tr\left(\rho
  P_{i}\right)$, where $P_{i}$ denotes the projector onto the
eigenspace corresponding to the eigenvalue $\lambda_{i}$ of the
Hermitian operator
$\mathbf{O}$~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.
Alternatively, one may say that the state $\rho$ determines the
probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as a function
from the set $\events$ of all possible projectors $P$ onto the
interval $[0,1]$, and satisfies the following probability axioms~
\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}:

\begin{itemize}
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

Furthermore, if we express an observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N}\lambda_{i}P_{i}$, the expectation
value of $\mathbf{O}$ is given by~\cite{peres1995quantum,544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\rho}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
\end{equation}

To formalize the idea of finite precision measurement, we next
introduce the notion of quantum interval-valued probability measures
(\emph{IVPMs})~$\bar{\mu}$. The latter replaces the infinitely precise
interval $[0,1]$ by a finite collection $\mathscr{I}$ of
intervals. Since the intervals $\imposs=\left[0,0\right]$ and
$\necess=\left[1,1\right]$ should be included in $\mathscr{I}$, and
arithmetic operations on intervals
satisfy \begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ .}
\end{cases}
\end{eqnarray}
\end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
defined by the axioms: 
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
condition that provides only a bound and not an equivalence as in
Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
$\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
$\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
\]

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Measurements}

@@

The above axioms describe a mathematical idealization involving
infinitely precise measurements. In an actual experimental setup with
an ensemble of quantum states that are not exactly identical, and with
small imperfections and inaccuracies in measuring devices, an
experimenter might only be able to determine that the probability of
the event $\proj{0}$ is in the range $[0.49,0.51]$. The spread in this
range depends on the amount of resources (time, energy, money, etc)
that are devoted to the experiment. In the classical setting, this
``fuzziness'' is formalized by moving to \emph{interval-valued
  probability measures}.

@@

Given a collection of intervals $\mathscr{I}$ with addition and scalar
multiplication defined as follows: 
\begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ ,}
\end{cases}
\end{eqnarray}
\end{subequations}
a QIVPM $\bar{\mu}$ is an assignment of intervals to each event
obeying the following axioms:
$\bar{\mu}(\mathbb{0})=\imposs=\left[0,0\right]$,
$\bar{\mu}(\mathbb{1})=\necess=\left[1,1\right]$,
$\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$,
and for a set of mutually orthogonal projections
$\left\{ P_{i}\right\} _{i=1}^{N}$, we have:
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
The first three axioms are the direct counterpart of the corresponding
axioms for conventional probability measures. The last condition
allows for the interval associated with a composite event to be
sharper than the interval calculated from each individual
event. Intuitively, consider an event $P$ and its complement
$\mathbb{1}-P$. Looking at each separately we may discover
$\bar{\mu}(P) = [0.2,0.3]$ and $\bar{\mu}(\mathbb{1}-P)=[0.7,0.8]$. By
adding the individual intervals, we predict the probability of the
composite event $P+(\mathbb{1}-P)$ to be in the range
$[0.9,1.1]$. Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}
allows us to refine this to $[1,1]$ by recognizing that the composite
event is the identity projection.

% In conventional quantum theory (CQT), when we measure a property
% $\mathbf{O}$ of a physical system whose state is represented by a
% density matrix~$\rho$, the probability to obtain the value $\lambda$
% of that property is determined by the Born rule
% $\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)=\Tr\left(\rho
%   P_{i}\right)$, where $P_{i}$ denotes the projector onto the
% eigenspace corresponding to the eigenvalue $\lambda_{i}$ of the
% Hermitian operator
% $\mathbf{O}$~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.
% Alternatively, one may say that the state $\rho$ determines the
% probability measure $\mu_{\rho}^{\mathrm{B}}$ defined as a function
% from the set $\events$ of all possible projectors $P$ onto the
% interval $[0,1]$, and satisfies the following probability axioms~
% \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}:

% \begin{itemize}
% \item $\mu_{\rho}^{\mathrm{B}}(\mathbb{0})=0$. 
% \item $\mu_{\rho}^{\mathrm{B}}(\mathbb{1})=1$. 
% \item $\mu_{\rho}^{\mathrm{B}}\left(\mathbb{1}-P\right)=
% 1-\mu_{\rho}^{\mathrm{B}}\left(P\right)$. 
% \item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
% $P_iP_j=\delta_{ij}P_i$,
% we have 
% \begin{equation}
% \mu_{\rho}^{\mathrm{B}}\left(\sum_{i=1}^{N}P_{i}\right)=
% \sum_{i=1}^{N}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)
% \textrm{ .}\label{eq:QuantumProbability-Addition}
% \end{equation}
% \end{itemize}
% Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
% satisfying the above axioms, Gleason's theorem asserts that there
% exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$
% in a Hilbert space of dimension $d\geq3$~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.

% Furthermore, if we express an observable~$\mathbf{O}$ as its spectrum
% decomposition~$\sum_{i=1}^{N}\lambda_{i}P_{i}$, the expectation
% value of $\mathbf{O}$ is given by~\cite{peres1995quantum,544199,Jaeger2007}
% \begin{equation}
% \expval{\mathbf{O}}_{\rho}=\Tr\left(\rho\mathbf{O}\right)=\sum_{i=1}^{N}\lambda_{i}\mu_{\rho}^{\mathrm{B}}\left(P_{i}\right)\ .\label{eq:quantum-expectation}
% \end{equation}

% To formalize the idea of finite precision measurement, we next
% introduce the notion of quantum interval-valued probability measures
% (\emph{IVPMs})~$\bar{\mu}$. The latter replaces the infinitely precise
% interval $[0,1]$ by a finite collection $\mathscr{I}$ of
% intervals. Since the intervals $\imposs=\left[0,0\right]$ and
% $\necess=\left[1,1\right]$ should be included in $\mathscr{I}$, and
% arithmetic operations on intervals
% satisfy \begin{subequations}\label{eq:interval-operations}
% \begin{eqnarray}
%  &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
%  &  & x[l,r]=\begin{cases}
% [xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
% [xr,xl] & \textrm{for }x\le0\textrm{ .}
% \end{cases}
% \end{eqnarray}
% \end{subequations} $\bar{\mu}:\events\rightarrow\mathscr{I}$ is
% defined by the axioms: 
% \begin{itemize}
% \item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
% \item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
% \item $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$. 
% \item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
% $P_{i}P_{j}=\delta_{ij}P_{i}$, we have 
% \begin{equation}
% \bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
% \end{equation}
% \end{itemize}
% condition that provides only a bound and not an equivalence as in
% Eq.~\eqref{eq:QuantumProbability-Addition}. For example, if we consider
% $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $ and
% $\bar{\mu}\left(P\right)=\left[l,r\right]$, we have, when $r-l=\ell>0$,
% \[
% \bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1-\ell,1+\ell\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right).
% \]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy States}
